
---
title: "Impacts of LOP differences on the net biodiversity effect<br> &nbsp;"
author: "Elena Plekhanova"
date: "`r format(Sys.time(), '%B %e, %Y')`"
output:
  html_document:
    df_print: paged
    code_folding: hide
  html_notebook:
    code_folding: hide
subtitle: <center> ---------------- 2D LOP experiment ---------------- </center>
fontsize: 14pt
---

<style type="text/css">

body{ /* Normal  */
      font-size: 14px;
  }
td {  /* Table  */
  font-size: 10px;
}
h1.title {
  font-size: 32px;
  color: DarkRed;
}

h1.subtitle {
  font-size: 32px;
  color: DarkRed;
}

h1 { /* Header 1 */
  font-size: 28px;
  color: DarkRed;
}
h2 { /* Header 2 */
    font-size: 23px;
  color: DarkRed;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

&nbsp;


We conducted an experiment, where we wanted to investigate how LOP of two "species" contribute to the net biodiversity effect. 
For this, we created around 10,000 simulations, where we vary:

* LOP of 1st "species"
  * reflectance of 1st "species" from 0 to 1 with interval 0.1
  * anglesittance of 1st "species" from 0 to (1 - reflectance) with interval 0.1
* LOP of 2nd "species"

For each combination of LOP, we created the scenes of monocultures and the scene of combination of this two species.

Other parameters of simulation:

* white Lambertian background
* number of trees = 4, repetitive scene
* no atmosphere
* direct sun, 0' zenith angle


```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(gridExtra)
library(gplots)
library(RColorBrewer)
library(cowplot)
library(viridis)

Nscenes = 1
Nspecies = 6
```

```{r}
solar_path = "./Data_for_figures/Solar_irradiance.csv"
read_solar_irradiance = function(sol_path = solar_path, bands = "all") {
  srad = read.csv(sol_path)
  srad = srad[srad$wavelength < 3,c(2:3)]
  srad$wavelength = round(srad$wavelength, 3)
  ## Pick only the irradiance between 400 and 2500 nm at 50nm range
  srad = srad[unlist(sapply(c(8:50)/20, function(i) which(abs(srad$wavelength-i)==min(abs(srad$wavelength-i))))),]
  srad = aggregate(srad, by = list(srad$wavelength), FUN = "mean")[,c(2:3)]
  if (bands == "VIS") {
  
  srad$irradiance = srad$irradiance/sum(srad$irradiance[1:7])
  srad$irradiance[8:43] = 0
  
  } else srad$irradiance = srad$irradiance/sum(srad$irradiance)
  
  return(srad$irradiance)
}

dat = read.csv("./Data_for_figures/50nm/profiles_exp1_lai_90.csv", stringsAsFactors = F)
dat = subset(dat, scenetype == 3 & height_m != 1 & num_of_species %in% c(1,2))

aggdat = aggregate(albedo ~ pos + num_of_species + band, data = dat, FUN = "mean")
  
solar_irradiance = read_solar_irradiance(bands = "all")
aggdat$fapar = (aggdat$albedo)*solar_irradiance[aggdat$band]

aggdat = aggregate(fapar ~ pos + num_of_species, data = aggdat, FUN = "sum")
aggdat$fapar = 1 - aggdat$fapar


cumsumm = cumsum(unlist(sapply(c(1:Nspecies), function(x) choose(Nspecies, x))))
the_list_of_comb = sapply(c(1:Nspecies), function(k) sapply(c(1:choose(Nspecies, k)), function(i) combn(Nspecies, k)[,i]))
the_list_of_sp = unlist(sapply(c(2:Nspecies), function(j) { x = the_list_of_comb[[j]]
res = split(x, rep(1:ncol(x), each = nrow(x)))
  return(res)}), recursive = FALSE)


diffs = lapply(c(2), function(k) sapply(c(1:choose(Nspecies, k)), function(i) {
  mix = aggdat$fapar[i+cumsumm[k-1]]
  av_mon = sum(aggdat$fapar[combn(Nspecies, k)[,i]])/k
  ratio = (mix - av_mon) / av_mon # calculating the gained radiation
  perc_ratio = ratio * 100 # in percent
  return(perc_ratio)
}))

diffs = as.data.frame(cbind(unlist(diffs), aggdat$num_of_species[-c(1:6)]))
colnames(diffs) = c("fapar_differences", "num_of_species")
diffs$pos = c(1:nrow(diffs))


lai_values = c(c(2:6), 1)
diffs$diff_lai_values = sapply(c(1:nrow(diffs)), function(i) abs(lai_values[the_list_of_sp[[i]][1]] - lai_values[the_list_of_sp[[i]][2]]))
```


```{r fig.height=3, fig.width=1.8}

my_palette = brewer.pal(5,"Greens")
my_palette=colorRampPalette(my_palette)(5)

ggplot(diffs, aes(diff_lai_values, fapar_differences, group = diff_lai_values, fill = as.factor(diff_lai_values))) +       
  geom_boxplot(width=0.5, alpha = 1, size = 0.65) + 
  scale_fill_manual(values = my_palette) + 
  xlab("LAI distance") + ylab("shortwave absorptance change, %") + theme_bw() +
  theme(panel.grid.minor.y = element_blank(), panel.grid.minor.x = element_blank(),
        plot.title = element_text(size=12, hjust = 0.5),
        axis.title.x = element_text(size=10), axis.text.x  = element_text(size=8),
        axis.title.y = element_text(size=10), axis.text.y  = element_text(size=8),
        legend.position = "none")
```

## Without vertical LAD

```{r message=FALSE, warning=FALSE}

subdat_crown$pos = subdat_crown$idSimulation
aggdat = aggregate(abs_mean ~ height_m + pos + angles + angles2 + num_of_species, 
                   data = subdat_crown, FUN = "mean")
aggdat = aggregate(abs_mean ~ pos + angles + angles2 + num_of_species, 
                   data = aggdat, FUN = "sum")



res_tab = lapply(unique(aggdat$pos[as.numeric(aggdat$num_of_species) > 1]), function(POS) {

  av_mixture = subset(aggdat, pos == POS, abs_mean)

  abs_of_monocultures = lapply(unique(subdat_crown$species[subdat_crown$pos == POS]), function(x) {
    position_of_mon = subdat_crown$pos[subdat_crown$species == x & 
                                         subdat_crown$num_of_species == 1][1]
    res = subset(aggdat, pos == position_of_mon, select = abs_mean)
    return(res)
  })
    
  av_monocultures = apply(do.call(rbind.data.frame, list(abs_of_monocultures)), 1, mean)
  
  one_mix = data.frame( ( sapply( c(1:nrow(av_mixture)), function(i) (av_mixture[i,1] - av_monocultures[i])*100/av_monocultures[i] ) ), c(1:nrow(av_mixture)), 
                        rep(POS, nrow(av_mixture)))
  colnames(one_mix) = c("abs_mean", "height_m", "pos")
  
  one_mix$diff_sum = rep(round(sum(one_mix$abs_mean), 7), nrow(one_mix))
  one_mix$num_of_species = rep(aggdat$num_of_species[aggdat$pos == POS][1], nrow(one_mix))
  one_mix$angles = rep(aggdat$angles[aggdat$pos == POS][1], nrow(one_mix))
  one_mix$angles2 = rep(aggdat$angles2[aggdat$pos == POS][1], nrow(one_mix))
  
  return(one_mix)
})
res_tab = do.call(rbind.data.frame, res_tab)

```


```{r fig.height=2.5, fig.width=3}
aggdat = aggregate(diff_sum ~ pos + angles + angles2, 
                   data = res_tab, FUN = "mean")
aggdat = aggdat[(aggdat$angles!= 0.0) & (aggdat$angles2!= 0.0),]

aggdat$diff_angles = abs(aggdat$angles - aggdat$angles2)
aggdat$diff_angles = as.factor(aggdat$diff_angles)
my_palette = brewer.pal(9,"Greens")
#my_palette = c("white", "#1B9E77")
my_palette=colorRampPalette(my_palette)(10)
mt = aggdat %>% mutate(diff_angles = diff_angles) %>%
  group_by(diff_angles) %>% 
  mutate(med_diffsum = median(diff_sum))
ggplot(mt, aes(diff_angles, diff_sum, group = diff_angles, fill = as.factor(med_diffsum))) +       geom_boxplot(width=0.5, alpha = 1, size = 0.65) + 
  scale_fill_manual(values = my_palette) + 
  xlab("Gap fraction distance") + ylab("absorption change, %") + theme_bw() +
  theme(panel.grid.minor.y = element_blank(),
        plot.title = element_text(size=12, hjust = 0.5),
        axis.title.x = element_text(size=10), axis.text.x  = element_text(size=8),
        axis.title.y = element_text(size=10), axis.text.y  = element_text(size=8),
        legend.position = "none")
```

## What combination of leaf optical properties contributes to this difference?

&nbsp;
H0: $abs_{mix-mon}$ mainly dependant on difference between absorptances of 2 species and their anglesittances.

To check if this sentance holds for our simulations, we represented difference between mixtures and monocultures in corresponding axes:

```{r fig.height=7, fig.width=7, message=FALSE, warning=FALSE}

aggdat = aggregate(diff_sum ~ pos + angles + angles2, 
                   data = res_tab, FUN = "mean")
aggdat$sp1 =  round( abs(aggdat$angles) , 2)
#aggdat$sp1 =  round( abs(aggdat$refl1 - aggdat$refl2) , 2)
aggdat$sp2 =  round( abs(aggdat$angles2) , 2)

vector_a = aggdat$sp1
vector_b = aggdat$sp2

albmat = sapply(sort(unique(vector_a)), function (a) 
  sapply(sort(unique(vector_b)), function(b) 
    mean( aggdat$diff_sum[vector_a == a & vector_b== b])))

my_palette = brewer.pal(7,"Greens")
my_palette=colorRampPalette(my_palette)(100)

lmat = rbind(c(0,3),c(2,1),c(0,4))
lwid = c(1.5,4)
lhei = c(1.5,4,1)


heatmap.2(t(albmat),margins=c(5,5), scale="none", labCol = sort(unique(vector_b)),
          labRow = sort(unique(vector_a)),
          dendrogram="none",  trace="none", xlab = "Difference in absorptance", key.title = NA, keysize = 1,
          lmat = lmat, lwid = lwid, lhei = lhei,
          key.xlab = "Absorption of mixtires - monocultures",
          ylab = "Difference in anglesittance", density.info="none", Colv = NA, Rowv = NA,
          col = my_palette, symm=F,symkey=F,symbreaks=F)

library(plotly)
p <- plot_ly(z = ~t(albmat)) %>% add_surface() %>%
  layout(
    title = "Difference in absorptance between mixtures and monocultures",
    scene = list(
      xaxis = list(title = "Difference in absorption"),
      yaxis = list(title = "Difference in anglesittance"),
      zaxis = list(title = "Difference")
    ))
#p

```

On natural background

## Optional

```{r}
brewer.pal(n = 8, name = "Dark2")
```


```{r fig.height=4, fig.width=8}


aggdat = aggregate(abs_mean ~ height_m + pos + angles + angles2 + num_of_species + species, 
                   data = subdat_crown, FUN = "mean")
res_tab = lapply(unique(aggdat$pos[as.numeric(aggdat$num_of_species) > 1]), function(POS) {

      av_mix = subset(aggdat, pos == POS)
      sps = unique(av_mix$species)
      av_mix_sp1 = av_mix$abs_mean[av_mix$species == sps[1]]
      av_mix_sp2 = av_mix$abs_mean[av_mix$species == sps[2]]
      av_mon_sp1 = subset(aggdat, species == sps[1] &  num_of_species == 1)$abs_mean
      av_mon_sp2 = subset(aggdat, species == sps[2] &  num_of_species == 1)$abs_mean

      one_mix = data.frame(cbind(c((av_mix_sp1 - av_mon_sp1)[1], (av_mix_sp2 - av_mon_sp2)[1]), 
                                 c(av_mon_sp1, av_mon_sp2),
                                 c(1, 0),
                                 rep(POS, 2) ))
      
      colnames(one_mix) = c("abs_mean", "mon_abs", "sp_a" ,"pos")
      one_mix$diff_mon = (av_mon_sp1 - av_mon_sp2)[1] ## difference in absorbtion of monocultures
      one_mix$diff_mix = (av_mix_sp1 - av_mix_sp2)[1] ## difference in absorbtion of mixtures
      one_mix$gain_mix = (av_mix_sp1 - av_mon_sp1)[1] + (av_mix_sp2 - av_mon_sp2)[1] ## difference in absorbtion of mixtures
      one_mix$species = sps
      one_mix$angles = av_mix$angles
      one_mix$angles2 = av_mix$angles2
      one_mix$name_species = paste0(sps[1], " + ", sps[2])
      
      return(one_mix)
    })

res_tab = do.call(rbind.data.frame, res_tab)

aggdat = aggregate(abs_mean ~ pos + angles + angles2 + sp_a, 
                   data = res_tab, FUN = "mean")
aggdat$maxangle = as.factor(sapply(c(1:nrow(aggdat)), function(i) max(aggdat$angles[i], aggdat$angles2[i])))

aggdat$diff_angles = abs(aggdat$angles - aggdat$angles2)
aggdat$diff_angles = as.factor(aggdat$diff_angles)
my_palette = brewer.pal(9,"Greens")
my_palette=colorRampPalette(my_palette)(10)
mt = aggdat %>% mutate(diff_angles = diff_angles) %>%
  group_by(diff_angles) %>% 
  mutate(med_diffsum = median(abs_mean))
ggplot(mt, aes(diff_angles, abs_mean, group = diff_angles, col = sp_a)) + geom_point() + 
  xlab("Difference in gap fraction") + ylab("Absorption of mix - mon") + theme_bw() +
  theme(panel.grid.minor.y = element_blank(),
        plot.title = element_text(size=12, hjust = 0.5),
        axis.title.x = element_text(size=10), axis.text.x  = element_text(size=8),
        axis.title.y = element_text(size=10), axis.text.y  = element_text(size=8))
```



```{r}
aggdat$maxangle = as.factor(sapply(c(1:nrow(aggdat)), function(i) max(aggdat$angles[i], aggdat$angles2[i])))

aggdat2 = aggregate(diff_sum ~ diff_angles + maxangle, 
                   data = aggdat, FUN = "mean")

ggplot(aggdat2, aes(diff_angles, diff_sum, group = diff_angles, col = maxangle)) + geom_point(size = 2) + geom_line(aes(group = maxangle)) + xlab("Difference in gap fraction") + ylab("FAPAR change") + theme_bw()
```


```{r fig.height=10, fig.width=16, message=FALSE, warning=FALSE}



heatmap.2(t(albmat),margins=c(7,7), scale="none", labCol = sort(unique(vector_b)),
          labRow = sort(unique(vector_a)),
          dendrogram="none",  trace="none", xlab = "Difference in absorption", key.title = NA, keysize = 1, 
          lmat = lmat, lwid = lwid, lhei = lhei,
          cexRow = 1.8, cexCol = 1.8,
          key.xlab = "FAPAR change", par = par(cex.lab = 1.8, cex.axis = 1.8),
          ylab = "Difference in anglesittance", density.info="none", Colv = NA, Rowv = NA,
          col = my_palette, symm=F,symkey=F,symbreaks=F)
par(cex.lab = 18)

```



&nbsp;
Although the most of variation indeed explained by these axes, we can see some non-linearity in this plot, which has not been explained yet.
&nbsp;